Ten Simple Rules for Teaching Programming

### 2017-07-07 paper.tex
Updating README

\citet{guzdial-myths} refers to the belief that some people are born programmers and others aren't
as ``computing's most enduring and damaging myth.''  This is often ``confirmed'' by looking at university grade distributions, which are commonly held to be bimodal: a low-scoring hump of those that will never get it, and a high-scoring hump of those that have the right stuff.
Our first and most important rule is that this is wrong:
competence at programming is not innate,
but is rather a learned skill that can be acquired and improved with practice.

@@

One-on-one tutoring is perhaps the ideal form of teaching: all of a teacher's attention can be focused on one student, and they can completely customise their teaching for their single student, and tailor individual feedback and corrections based on two-way dialogue with that student.  In realistic settings however, one teacher must teach many students at once, whether that is a handful, tens or even hundreds in large lectures.  In larger settings, how can teachers possibly hope to clear up learners' many different misconceptions in a reasonable time?

@@

\begin{figure}\sffamily
\begin{verbatim}
for (int i = 1; i < 10; i++)
{
     if (i < 3 || i > 8)
     {
          System.out.println("Yes");
     }
}
\end{verbatim}
How many times will the above code print out the word Yes?\\
a) 10\\
b) 9\\
c) 4\\
d) 3
\label{fig:mcq}
\caption{An example multiple choice question, probing learners' understanding of loops and integer comparisons.}
\end{figure}

@@

\item It facilitates unintended knowledge transfer: students learn more
  than the instructor consciously intends to teach by watching \textit{how}
  instructors do things.  The extra knowledge may be high-level (e.g. whether a program is written top-down or bottom-up) or fairly low-level (e.g. using keyboard shortcuts).

@@

\item Instructors can spend too much time typing in boilerplate code that is needed by the lesson, but not
  directly relevant to it (such as library import statements), can
  distract learners from the intended thrust of a lesson.  As
  \citet{willingham-book} says, ``Memory is the residue of thought'';
  if the instructor spends their time typing boilerplate, that may be
  all that learners take away.  This can be avoided by starting with a partial skeleton that includes the boilerplate, or having it on hand to copy and paste at the appropriate juncture.

@@

When instructors are using live coding,
they usually run the program several times during its development to show what it does.
Research from peer instruction in physics education shows that
learners who observe a demonstration \textit{do not} learn better than
those who did not see the demonstration~\citep{crouch-demo},
and in fact many learners misremember the outcome of demonstrations afterwards~\citep{miller-demo}.  That is, demonstrations can be useless or actively harmful to learning.

>>>>

\citet{guzdial-myths} refers to the belief that some people are born programmers and others aren't
as ``computing's most enduring and damaging myth.''
This is often ``confirmed'' by looking at university grade distributions,
which are commonly held to be bimodal:
a low-scoring hump of those that will never get it
and a high-scoring hump of those that have the right stuff.
Our first and most important rule is that this is wrong:
competence at programming is not innate,
but is rather a learned skill that can be acquired and improved with practice.

@@

One-on-one tutoring is the ideal form of teaching:
all of a teacher's attention can be focused on one student,
and they can completely customise their teaching for that person
and tailor individual feedback and corrections based on two-way dialogue with them.
In realistic settings,
however,
one teacher must teach a handful, tens, or hundreds of students at once.
How can teachers possibly hope to clear up many learners' different misconceptions
in these larger settings in a reasonable time?

@@

\begin{figure}\sffamily
\begin{verbatim}
for (int i = 1; i < 10; i++) {
    if (i < 3 || i > 8) {
        System.out.println("Yes");
    }
}
\end{verbatim}
How many times will the above code print out the word Yes?\\
a) 10\\
b) 9\\
c) 4\\
d) 3
\label{fig:mcq}
\caption{An example multiple choice question probing learners' understanding of loops and integer comparisons.}
\end{figure}

@@

\item It facilitates unintended knowledge transfer: students learn
  more than the instructor consciously intends to teach by watching
  \textit{how} instructors do things.  The extra knowledge may be
  high-level (e.g., whether a program is written top-down or
  bottom-up) or fairly low-level (e.g., using keyboard shortcuts).

@@

\item Instructors can spend too much time typing in boilerplate code
  that is needed by the lesson, but not directly relevant to it (such
  as library import statements), can distract learners from the
  intended thrust of a lesson.  As \citet{willingham-book} says,
  ``Memory is the residue of thought''; if the instructor spends their
  time typing boilerplate, that may be all that learners take away.
  This can be avoided by starting with a partial skeleton that
  includes the boilerplate, or having it on hand to copy and paste
  when needed.  (Of the two, we prefer the former, since learners
  may not be able to keep up with copying and pasting.)

@@

When instructors are using live coding, they usually run the program
several times during its development to show what it does.  Research
from peer instruction in physics education shows that learners who
observe a demonstration \textit{do not} learn better than those who
did not see the demonstration~\citep{crouch-demo}, and in fact many
learners misremember the outcome of demonstrations
afterwards~\citep{miller-demo}.  In other words, demonstrations can
actually be useless or actively harmful.
